### 2019-07-28 reports/thesis/chapters/3_compoundPoisson.tex
Revise compound Poisson chapter

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.)
\begin{equation}
  \prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!} \quad \text{for }y=0,1,2,\cdots
\end{equation}
where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\dotdotdot$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ with probability $\prob(X=0)=\euler^{-\lambda}$. Then $X$ has probability mass at $X=0$ and probability density at $X>0$ which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

@@

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a a grey value. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. In a quantum counter, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value is obtained.

An energy integrating detector records the grey value which is linear to the energy detected \citep{whiting2006properties}. The grey value $X$ is
\begin{equation}
X|Y = \sum_{i=1}^Y U_i + \epsilon \ .
\end{equation}
This is the compound Poisson with Normal noise added to it. The scale factor $b$ is not included as this can be absorbed into $U$. Using the result that $\expectation\left[X\right]=\expectation\expectation\left[X|Y\right]$ and $\variance\left[X\right] = \variance\expectation\left[X|Y\right] + \expectation\variance\left[X|Y\right]$, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = \lambda\expectation\left[U\right]+a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = \lambda \expectation\left[U^2\right]+\kappa
\end{equation}
respectively. Eliminating $\lambda$ obtains
\begin{equation}
\variance\left[X\right] = \dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \expectation\left[X\right] + \kappa - a\dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \ .
\end{equation}
By assuming no beam hardening, $\expectation\left[U\right]$ and $\expectation\left[U^2\right]$ remains constant, thus there is a linear relationship between the variance and expectation of the grey value. There are other types of detection schemes \citep{whiting2006properties} but it shall be not be considered here.

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethylene cylinder \citep{yang2009evaluation} \citep{yang2010noise} at different voltages and powers. It was found there were 2 components in the noise, one was signal dependent and comes from the compound Poisson, the other was signal independent and may be electronic noise. The electronic noise can be modelled as Normally distributed \citep{xu2009electronic}.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be
\begin{equation}
  M_X(\theta)=\expectation\left[\euler^{X\theta}\right]
  \ .
\end{equation}
This can be computed using the result for conditional expectations $M_X(\theta)=\expectation\expectation\left[\euler^{X\theta}|Y\right]$. Using the definition of $X|Y$ in Equation \eqref{eq:compoundPoisson_X|Y}, then
\begin{align}
  M_X(\theta)&=\expectation\expectation\left[\exp\left(\theta U_1 + \theta U_2 + \dotdotdot + \theta U_Y\right)|Y\right]
  \\
  &=\expectation\expectation\left[\euler^{\theta U_1}\cdot\euler^{\theta U_2}\cdot\dotdotdot\cdot\euler^{\theta U_Y}|Y\right]
  \ .
\end{align}
But because $U_i$ for $i=1,2,3,\dotdotdot$ are i.i.d., then each $U_i$ has a common m.g.f.~$M_U(\theta)=\expectation\left[\euler^{U\theta}\right]$, then
\begin{align}
  M_X(\theta)&=\expectation\left(
    \expectation\left[\euler^{\theta U_1}|Y\right]\cdot
    \expectation\left[\euler^{\theta U_2}|Y\right]\cdot
    \dotdotdot \cdot
    \expectation\left[\euler^{\theta U_Y}|Y\right]
  \right)
  \\
  &=\expectation\left[\left(M_U(\theta)\right)^Y\right]
  \\
  &=\expectation\left[\euler^{Y\ln(M_U(\theta))}\right]
  \\
  & = M_Y\left(\ln(M_U(\theta)\right)
\end{align}
where $M_Y(\theta)$ is the m.g.f.~of $Y$. It can be shown that the m.g.f.~of $Y$ is
$
  M_Y(\theta)=\expectation\left[\euler^{Y\theta}\right]=
  \exp
  \left[
    \lambda
    \left(
      \euler^\theta-1
    \right)
  \right]
$
then
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
  \ .
\end{equation}

Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting it to zero
\begin{align}
  M_X'(\theta)&=\exp\left[\lambda\left(M_U(\theta)-1\right)\right]\cdot\lambda M_U'(\theta) \\
  &=M_X(\theta)\lambda M_U'(\theta)
\end{align}
then
\begin{equation}
  \expectation\left[X\right]=\lambda\expectation\left[U\right]
  \ .
\end{equation}
Conducting the same procedure
\begin{equation}
  M_X''(\theta)=M_X'(\theta)\lambda M_U'(\theta)+M_X(\theta)\lambda M_U''(\theta)
  \ ,
\end{equation}
the variance can be obtained
\begin{align}
  \variance\left[X\right]&=M_X''(0)-\left[M_X'(0)\right]^2
  \\
  &=M_X'(0)\lambda M_U'(0)+M_X(0)\lambda M_U''(0)-\left[\lambda\expectation\left[U\right]\right]^2
  \\
  &=\lambda^2 \left(\expectation\left[U\right]\right)^2+\lambda \expectation\left[U^2\right]-\left[\lambda\expectation\left[U\right]\right]^2
\end{align}
resulting in
\begin{equation}
  \variance\left[X\right] = \lambda\expectation\left[U^2\right]
  \ .
\end{equation}
This can be extended for higher moments.

Because the m.g.f.~can be written down in closed form, so is the characterisation function or Fourier transform of the p.d.f.~of $X$. By obtaining an empirical version of $p_U(u)$, moments of $X$ can be estimated using the m.g.f.~and the p.d.f.~can be estimated by using fast Fourier transform on the empirical characteristic function \citep{whiting2006properties}.

@@

A special case of the compound Poisson is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. The special case is used for example in \cite{xu2009electronic}. This is known as the compound Poisson-gamma distribution and denoted as $X\sim\CPoisson(\lambda,\alpha,\beta)$ and has p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \sum_{y=1}^{\infty}\frac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
The conditional distribution can be shown to be $X|Y\sim\gammaDist\left(Y\alpha,\beta\right)$.

It can be shown that the compound Poisson-gamma distribution is in the exponential family for fixed $\alpha$ \citep{jorgensen1987exponential}. To show this, it is useful to reparametrize the compound Poisson-gamma distribution using the following:
\begin{equation}
  p=\frac{2+\alpha}{1+\alpha}
  \ ,
\end{equation}
\begin{equation}
  \mu=\frac{\lambda\alpha}{\beta}
  \ ,
\end{equation}
\begin{equation}
  \phi = \frac{\alpha+1}{\beta^{2-p}(\lambda\alpha)^{p-1}}
  \ .
\end{equation}
The parameters $p$, $\mu$ and $\phi$ are called the index, mean and dispersion parameters respectively. It can be shown that $1<p<2$, $\mu>0$ and $\phi>0$.

By rearranging the parameters to get
\begin{equation}
  \lambda=\frac{\mu^{2-p}}{\phi(2-p)}
\end{equation}
\begin{equation}
  \alpha=\frac{2-p}{p-1}
\end{equation}
\begin{equation}
  \beta=\frac{1}{\phi(p-1)\mu^{p-1}}
\end{equation}
and substituting it into Equation \eqref{eq:compoundPoisson_pdf}, the p.m.f.~at zero can be shown to be
\begin{equation}
  \prob(X=0) = \exp
  \left[
      -\frac{\mu^{2-p}}{\phi(2-p)}
  \right]
\end{equation}
and the p.d.f.~for $x>0$ is
\begin{multline}
  p_X(x) = \sum_{y=1}^{\infty}
  \left[
    \frac{1}{\phi(p-1)\mu^{p-1}}
  \right]^{y\alpha}
  \frac{1}{\Gamma(y\alpha)}
  x^{y\alpha-1}
  \exp\left[
      -\frac{x}{\phi(p-1)\mu^{p-1}}
  \right]
  \\
  \exp\left[
      -\frac{\mu^{2-p}}{\phi(2-p)}
  \right]
  \left[
    \frac{\mu^{2-p}}{\phi(2-p)}
  \right]^y
  \frac{1}{y!}
  \ .
\end{multline}
Tidying up the equation
\begin{multline}
  p_X(x) = 
  \exp\left[
    \frac{1}{\phi}\left(x\frac{\mu^{1-p}}{1-p}-\frac{\mu^{2-p}}{2-p}\right)
  \right]
  \frac{1}{x}
  \\
  \sum_{y=1}^{\infty}\frac{x^{y\alpha}\mu^{y[2-p-\alpha(p-1)]}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
  \ .
\end{multline}
To simplify further, it should be noted that $2-p-\alpha(p-1) = 2-p - \frac{2-p}{p-1}(p-1) =0$ so that
\begin{equation}
  p_X(x) = 
  \exp\left[
    \frac{1}{\phi}
    \left(
      x\frac{\mu^{1-p}}{1-p}-\frac{\mu^{2-p}}{2-p}
    \right)
  \right]
  \frac{1}{x}
  \sum_{y=1}^{\infty}W_y(x,p,\phi)
\end{equation}
where
\begin{equation}
  W_y(x,p,\phi)=\frac{x^{y\alpha}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
  \ .
\end{equation}
This is in the form of a generalised linear model \citep{nelder1972generalized, nelder1972generalized_2, mccullagh1984generalized} for fixed $p$ or $\alpha$ because the above is in the form of a distribution in the dispersive exponential family. Parameter estimation then can be done via the generalised linear model framework and can be extended to include linear mixed models \citep{zhang2013likelihood}. These has applications in for example insurance claim data \citep{jorgensen1994fitting, smyth2002fitting}.

Estimating $p$ is difficult and various methods were discussed by \cite{zhang2013likelihood}. One easy way is to estimate $\mu$ and $\phi$ on a grid of $p$'s and then select the $p$ which maximises the likelihood \citep{dunn2005series}.

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It has a special variance mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where, as a reminder, $1<p<2$. This can be derived using the partition function $Z$. Let
\begin{equation}
  \theta = \frac{\mu^{1-p}}{1-p}
\end{equation}
and
\begin{equation}
  \ln Z = \frac{1}{\phi(2-p)\theta^\alpha(1-p)^\alpha} \ ,
\end{equation}
then
\begin{equation}
  \variance[X] = \phi^2 \frac{\partial^2\ln Z}{\partial\theta^2} \ .
\end{equation}

@@

The method of moments is a simpler method to estimate the parameters of a compound Poisson-gamma random variable. The m.g.f.~of the of $X$ is
\begin{equation}
  M_X(\theta)=\exp\left[\lambda\left(\left(\frac{\beta}{\beta-\theta}\right)^{\alpha}-1\right)\right]
\end{equation}
and moments can be obtained from it such as
\begin{equation}
  \expectation\left[X\right]=\frac{\alpha\lambda}{\beta}
\end{equation}
\begin{equation}
  \variance\left[X\right]=\frac{\alpha(\alpha+1)\lambda}{\beta^2}
  \label{eq:compoundPoisson_variance}
\end{equation}
and
\begin{equation}
  \expectation\left[(X-\expectation[X])^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\lambda}{\beta^3}
  \ .
\end{equation}

Suppose $\widehat{\mu}$ is an estimator of $\expectation[X]$ and $\widehat{\mu}_j$ is an estimator of $\expectation\left[\left(X-\expectation[X]\right)^j\right]$ for $j=2,3$. Then the estimators
\begin{equation}
  \widehat{\lambda}=\frac{\widehat{\mu}^2\widehat{\mu}_2}{\left(2\widehat{\mu}_2^2-\widehat{\mu}_3\widehat{\mu}\right)}
\end{equation}
\begin{equation}
  \widehat{\alpha}=\frac{\widehat{\mu}_3\widehat{\mu}-2\widehat{\mu}_2^2}{\widehat{\mu}_2^2-\widehat{\mu}\widehat{\mu}_3}
\end{equation}
\begin{equation}
  \widehat{\beta}=\frac{\widehat{\mu}\widehat{\mu}_2}{\widehat{\mu}\widehat{\mu}_3-\widehat{\mu}_2^2}
\end{equation}
are method of moments estimators of $\lambda$, $\alpha$ and $\beta$ respectively \citep{withers2011compound}. These estimators suffer because estimation is not done through the sufficient statistics and can be negative, this is a problem because the parameters do not take non-positive values.

@@

The m.g.f.~of $X$ is
\begin{equation}
M(\theta)=\exp\left[\lambda\left(\left(\frac{\beta}{\beta-\theta}\right)^{\alpha}-1\right)\right]
\ .
\end{equation}
It provides a starting point to what limiting distributions the compound Poisson-gamma distribution converges to for large parameters. These limiting distributions then can be used to approximate the p.d.f.~of a compound Poisson-Gamma random variable. The m.g.f.~of $X$ itself it not useful because when considering $\lambda\rightarrow\infty$ or $\alpha\rightarrow\infty$ then $M_X(\theta)\rightarrow\infty$. Also for $\beta\rightarrow\infty$, then $M_X(\theta)\rightarrow 1$ which is not useful either.

The compound Poisson-gamma random variable $X$ can be standardised to obtain useful limiting results from the m.g.f. Let
\begin{equation}
  Z = \frac{X-\expectation[X]}{\sqrt{\variance[X]}}
  \ ,
\end{equation}
then using previous results
\begin{equation}
  Z = bX+a
\end{equation}
where
\begin{equation}
  b = \frac{\beta}{\sqrt{\alpha(\alpha+1)\lambda}}
\end{equation}
and
\begin{equation}
  a = -\sqrt{\frac{\alpha\lambda}{\alpha+1}}
  \ .
\end{equation}

The m.g.f.~of $Z$ is then
\begin{align}
  M_Z(\theta)&=\expectation\left[\euler^{Z\theta}\right]
  \\
  &=\expectation\left[\euler^{(bX+a)\theta}\right]
  \\
  &=\euler^{a\theta}M_X(b\theta)
  \ .
\end{align}
Substituting in the values
\begin{align}
  M_Z(\theta)&=
  \exp\left(
      -\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
    \right)
  \exp\left[
    \lambda
    \left(
      \left(
        \frac{\beta}{\beta-\frac{\beta\theta}{\sqrt{\alpha(\alpha+1)\lambda}}}
      \right)^\alpha
      -1
    \right)
  \right]
  \\
  &=
  \exp\left(
      -\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
    \right)
  \exp\left[
    \lambda
    \left(
      \left(
        \frac{\sqrt{\alpha(\alpha+1)\lambda}}{\sqrt{\alpha(\alpha+1)\lambda}-\theta}
      \right)^\alpha
      -1
    \right)
  \right]
\end{align}
or in a different form
\begin{equation}
	M_Z(\theta)=
	\exp\left(
	    -\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
    \right)
	\exp\left[
		\lambda
		\left(
			\left(
				1-\frac{\theta}{\sqrt{\alpha(\alpha+1)\lambda}}	
			\right)^{-\alpha}
			-1
		\right)
	\right]
	\ .
\end{equation}

The above form is useful so that the binomial expansion can be conducted, that is
\begin{multline}
	\left(
		1-\frac{\theta}{\sqrt{\alpha(\alpha+1)\lambda}}
	\right)^{-\alpha}
	=
	1+
	\sum_{r=1}^\infty \frac{\theta^r\prod_{s=1}^r(\alpha+s-1)}{(\alpha(\alpha+1)\lambda)^{r/2}r!}
	\\
	\text{for }\frac{\|\theta\|}{\sqrt{\alpha(\alpha+1)\lambda}}<1 \ .
\end{multline}
Substituting in the binomial expansion
\begin{equation}
	M_Z(\theta)=
	\exp\left(
	    -\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
    \right)
	\exp\left[
		\lambda
		\sum_{r=1}^\infty \frac{\theta^r\prod_{s=1}^r(\alpha+s-1)}{(\alpha(\alpha+1)\lambda)^{r/2}r!}
	\right]
	\ .
\end{equation}
Writing in full the $r=1,2$ terms
\begin{multline}
	M_Z(\theta)=
	\exp\left(
	    -\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
    \right)
	\exp\left[
		\theta\sqrt{\frac{\alpha\lambda}{\alpha+1}}
		+\frac{\theta^2}{2}
	\right.
	\\
	\left.
		+\lambda\sum_{r=3}^\infty \frac{\theta^r\prod_{s=1}^r(\alpha+s-1)}{(\alpha(\alpha+1)\lambda)^{r/2}r!}
	\right]
\end{multline}
and a term cancels out to get
\begin{equation}
	M_Z(\theta)=
	\exp\left[
		\frac{\theta^2}{2}
		+\sum_{r=3}^\infty
		\frac
			{\theta^r\prod_{s=1}^r(\alpha+s-1)}
			{(\alpha(\alpha+1))^{r/2}r!}
		\lambda^{1-r/2}
	\right]
	\ .
\end{equation}

For large $\lambda$
\begin{equation}
	\lim_{\lambda\rightarrow\infty}M_Z(\theta)=\exp\left[\frac{\theta^2}{2}\right]
\end{equation}
which is the same as the m.g.f.~of a standard Normal variable, therefore
\begin{equation}
	\lim_{\lambda\rightarrow\infty}Z\sim\normal(0,1) \ .
\end{equation}
This should make sense as for high $\lambda$, the Poisson random variable has a high expectation, increasing the number of gamma random variables in a summation. Increasing the number of terms in a summation will trigger the central limit theorem.

Next for high $\alpha$, it should be noted that
\begin{equation}
  \lim_{\alpha\rightarrow\infty}
  \frac{
    \prod_{s=1}^r(\alpha+s-1)
  }
  {
    \alpha(\alpha+1))^{r/2}
  }
  =
  \frac{
    \prod_{s=1}^r\alpha
  }
  {
    \alpha^{r}
  }
\end{equation}
so that
\begin{equation}
  \lim_{\alpha\rightarrow\infty}
  \frac{
    \prod_{s=1}^r(\alpha+s-1)
  }
  {
    (\alpha(\alpha+1))^{r/2}
  }
  = 1
  \ .
\end{equation}
As a result
\begin{equation}
  \lim_{\alpha\rightarrow\infty}
  M_Z(\theta)=
  \exp\left[
    \frac{\theta^2}{2}
    +\sum_{r=3}^\infty
    \frac{\theta^r}{r!}
    \lambda^{1-r/2}
  \right]
  \ .
\end{equation}
However this shows that taking the limit $\alpha\rightarrow\infty$ is not enough to get a Normal limiting distribution. For a Normal limiting distribution, the limit must be accompanied with the limit $\lambda\rightarrow\infty$, that is
\begin{equation}
  \lim_{\lambda\rightarrow\infty}\lim_{\alpha\rightarrow\infty}Z\sim\normal(0,1)
  \ .
\end{equation}

Finally it should be noted that $M_Z(\theta)$ is independent of $\beta$. Thus $\beta$ will have no effect on the convergence to a Normal limiting distribution.

The above results justify the use of the approximation
\begin{equation}
	X\sim\normal\left(\frac{\lambda\alpha}{\beta},\frac{\lambda\alpha(\alpha+1)}{\beta^2}\right)
\end{equation}
for large $\lambda$. The limiting case where $\lambda\rightarrow 0$, $\alpha\rightarrow 0$ and $\beta\rightarrow 0$ will not be discussed here.

@@

Simulations of a compound Poisson-gamma random variable were conducted with the aim to compare how well these density evaluations methods perform. This was done by comparing the evaluated densities with the histogram of simulations and a Q-Q plot. The following compound Poisson-gamma distributions were used in the simulations to capture the variety in the compound Poisson-gamma family: $\CPoisson(1,1,1)$, $\CPoisson(10,1,1)$, $\CPoisson(1,100,1)$, $\CPoisson(100,100,1)$.

>>>>

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.) $\prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!}$ for $y=0,1,2,\cdots$, where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\cdots$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ with probability $\prob(X=0)=\euler^{-\lambda}$. Then $X$ has probability mass at zero and probability density for positive numbers which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

@@

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a digital signal. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. In a quantum counter, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value is obtained.

An energy integrating detector records the grey value as linear to the energy detected \citep{whiting2006properties}. The grey value $X$ is
\begin{equation}
X|Y = \sum_{i=1}^Y U_i + \epsilon \ .
\end{equation}
This is the compound Poisson with Normal noise added to it. The scale factor $b$ is not included as this can be absorbed into $U$. Using the result that $\expectation\left[X\right]=\expectation\expectation\left[X|Y\right]$ and $\variance\left[X\right] = \variance\expectation\left[X|Y\right] + \expectation\variance\left[X|Y\right]$, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = \lambda\expectation\left[U\right]+a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = \lambda \expectation\left[U^2\right]+\kappa
\end{equation}
respectively. Eliminating $\lambda$ obtains
\begin{equation}
\variance\left[X\right] = \dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \expectation\left[X\right] + \kappa - a\dfrac{\expectation\left[U^2\right]}{\expectation\left[U\right]} \ .
\end{equation}
By assuming no beam hardening, $\expectation\left[U\right]$ and $\expectation\left[U^2\right]$ remains constant, thus there is a linear relationship between the variance and expectation of the grey value. There are other types of detection schemes \citep{whiting2006properties} but it shall be not be considered here.

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethylene cylinder \citep{yang2009evaluation, yang2010noise} at different voltages and powers. It was found there were 2 components in the noise, one was signal dependent and comes from the compound Poisson, the other was signal independent and may be electronic noise. The electronic noise can be modelled as Normally distributed \citep{xu2009electronic}.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right] \ .
\end{equation}
The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting it to zero. In other words, $\expectation[X^r]=M_X^{(r)}(\theta)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
\end{align}
which agrees with the results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

\subsection{Compound Poisson-Gamma}

A special case of the compound Poisson is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. This was used for example in \cite{xu2009electronic}. This is known as the compound Poisson-gamma distribution. This is denoted by $X\sim\CPoisson(\lambda,\alpha,\beta)$ and has p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \sum_{y=1}^{\infty}\frac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
Recall that $X|Y=\sum_{i=1}^YU_i$ which involves a sum of gamma random variables. It can be shown that
\begin{equation}
  X|Y\sim\gammaDist\left(Y\alpha,\beta\right).
\end{equation}

The m.g.f.~of $U$ is $M_U(\theta) = \left(\dfrac{\beta}{\beta-\alpha}\right)^\alpha$. Then the m.g.f.~of $X$ is
\begin{equation}
  M_X(\theta)=\exp\left[\lambda\left(\left(\frac{\beta}{\beta-\theta}\right)^{\alpha}-1\right)\right]
\end{equation}
and moments can be obtained from it such as
\begin{equation}
  \expectation\left[X\right]=\frac{\alpha\lambda}{\beta}
\end{equation}
\begin{equation}
  \variance\left[X\right]=\frac{\alpha(\alpha+1)\lambda}{\beta^2}
  \label{eq:compoundPoisson_variance}
\end{equation}
and
\begin{equation}
  \expectation\left[(X-\expectation[X])^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\lambda}{\beta^3}
  \ .
\end{equation}

@@

It can be shown that the compound Poisson-gamma distribution is in the exponential family for fixed $\alpha$ \citep{jorgensen1987exponential}. To show this, it is useful to parametrise the compound Poisson-gamma distribution using the following:
\begin{equation}
  p=\frac{2+\alpha}{1+\alpha}
  \ ,
\end{equation}
\begin{equation}
  \mu=\frac{\lambda\alpha}{\beta}
  \ ,
\end{equation}
\begin{equation}
  \phi = \frac{\alpha+1}{\beta^{2-p}(\lambda\alpha)^{p-1}}
  \ .
\end{equation}
The parameters $p$, $\mu$ and $\phi$ are called the index, mean and dispersion parameters respectively. The parameters take the values of $1<p<2$, $\mu>0$ and $\phi>0$. After parametrising, it can be shown that the p.m.f.~at zero is
\begin{equation}
  \prob(X=0) = \exp
  \left[
      -\frac{\mu^{2-p}}{\phi(2-p)}
  \right]
\end{equation}
and the p.d.f.~for $x>0$ is
\begin{equation}
  p_X(x) = 
  \exp\left[
    \frac{1}{\phi}
    \left(
      x\frac{\mu^{1-p}}{1-p}-\frac{\mu^{2-p}}{2-p}
    \right)
  \right]
  \frac{1}{x}
  \sum_{y=1}^{\infty}W_y(x,p,\phi)
\end{equation}
where
\begin{equation}
  W_y = W_y(x,p,\phi)=\frac{x^{y\alpha}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
  \ .
\end{equation}
The derivation is shown in Appendix \ref{chapter:appendix_tweedie}. This is in the form of a generalised linear model \citep{nelder1972generalized, nelder1972generalized_2, mccullagh1984generalized} for fixed $p$ or $\alpha$ because the above is in the form of a distribution in the dispersive exponential family. This has applications in, for example, insurance claim data \citep{jorgensen1994fitting, smyth2002fitting}.

Parameter estimation for fixed $p$ can be done via the generalised linear model framework and can be extended to include linear mixed models \citep{zhang2013likelihood}. Estimating $p$ is difficult and various methods were discussed by \cite{zhang2013likelihood}. One way is to estimate $\mu$ and $\phi$ on a grid of $p$'s and then select the $p$ which maximises the likelihood \citep{dunn2005series}.

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It can be shown that it has a special variance mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where $1<p<2$. This is derived in Appendix \ref{chapter:appendix_tweedie}. It should be noted that this relationship is for fixed $p$. This is different from the linear variance mean relationship found at the start of the chapter which was for fixed $\alpha$ from assuming no beam hardening.

@@

The method of moments is a simpler method to estimate the parameters of a compound Poisson-gamma random variable. Suppose $\widehat{\mu}$ is an estimator of $\expectation[X]$ and $\widehat{\mu}_j$ is an estimator of $\expectation\left[\left(X-\expectation[X]\right)^j\right]$ for $j=2,3$. Then the estimators
\begin{equation}
  \widehat{\lambda}=\frac{\widehat{\mu}^2\widehat{\mu}_2}{\left(2\widehat{\mu}_2^2-\widehat{\mu}_3\widehat{\mu}\right)}
\end{equation}
\begin{equation}
  \widehat{\alpha}=\frac{\widehat{\mu}_3\widehat{\mu}-2\widehat{\mu}_2^2}{\widehat{\mu}_2^2-\widehat{\mu}\widehat{\mu}_3}
\end{equation}
\begin{equation}
  \widehat{\beta}=\frac{\widehat{\mu}\widehat{\mu}_2}{\widehat{\mu}\widehat{\mu}_3-\widehat{\mu}_2^2}
\end{equation}
are method of moments estimators of $\lambda$, $\alpha$ and $\beta$ respectively \citep{withers2011compound}. These estimators suffer because estimation is not done through the sufficient statistics and can be negative, this is a problem because the parameters do not take non-positive values.

@@

The m.g.f.~provides a starting point to what limiting distributions the compound Poisson-gamma distribution converges to for large parameters. These limiting distributions can be used to approximate the p.d.f.~of the compound Poisson-gamma distribution.

It can be shown for large $\lambda$, the Normal approximation is
\begin{equation}
  X\sim\normal\left(\frac{\lambda\alpha}{\beta},\frac{\lambda\alpha(\alpha+1)}{\beta^2}\right) \ .
\end{equation}
The solution is shown in Appendix \ref{chapter:appendix_normalApproximation}.

@@

Simulations of a compound Poisson-gamma random variable were conducted with the aim to compare how well these density evaluations methods perform. This was done by comparing the evaluated densities with the histogram of simulations and a Q-Q plot. The following compound Poisson-gamma distributions were used in the simulations to capture the variety in the compound Poisson-gamma family: $\CPoisson(1,1,1)$, $\CPoisson(1,100,1)$, $\CPoisson(10,1,1)$, $\CPoisson(100,100,1)$. Varying $\beta$ is not interesting here as this only scales the random variable.
